Forecasting the old-age dependency ratio to determine the best pension age

### 2020-09-15 PensionAge.Rmd
Revised Section 3

```{r changespension, echo=FALSE, fig.cap="Government's adjustments on pension age scheme.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Policies obtained from https://www.dss.gov.au/seniors/benefits-payments/age-pension
# Law: https://guides.dss.gov.au/guide-social-security-law/3/4/1/10
# History: https://nationalseniors.com.au/news/campaign/a-short-history-of-the-age-pension-section
pension_age <- tibble(
  Year = 1920:2068,
  Current = 65
) %>%
  mutate(
    Proposed = pmin(70, Current + 0.5 * pmax(0, trunc((Year - 2015) / 2))),
    Approved = pmin(Proposed, 67),
  ) %>%
  pivot_longer(-Year, values_to = "Age", names_to = "Policy")

pension_age %>%
  filter(Year >= 2008, Year <= 2050) %>%
  ggplot(aes(x = Year, y = Age, group = Policy, col = Policy)) +
  geom_line() +
  ylab("Age of retirement") +
  scale_color_brewer(
    palette = "Dark2",
    breaks = c("Proposed", "Approved", "Current"),
    labels = c("Proposed further change", "Approved change", "Pension age at 65")
  )
```

@@

Variants of the Lee-Carter method [@LC92] are frequently used for demographic forecasting [see, e.g., @SBH11]. The most recent forecast of the population in Australia by the @PC13 was based on basic Lee-Carter methods plus some judgement-based adaptations to address the characteristics of demographic trends that the Lee-Carter method failed to capture. The approach of forecasting used in this study is similar to the @PC13. The Hyndman-Ullah method (2007) is used as a basic forecasting approach, which is a comprehensive extension of the Lee-Carter method. In addition, an adapted product-ratio method [@HBY13] is used to eliminate the potential long-term divergence between female and male mortality rates.

@@

The model is over-parameterised. Therefore, the constraints
\begin{equation}
  \sum_{t=1}^{n}\beta_t=0, \qquad\text{and}\qquad \int_x \phi(x)=1
\end{equation}
are imposed, and a singular value decomposition is used to find the least squares solution of \eqref{eq:lc} [@Good1969]. The estimates of $\phi(x)$ and $\beta_t$ are therefore the first principal component and scores of $\bm{X}$ with $(i,t)$th element $[\log(m_{t}(x))-\mu(x)]$, where $\mu(x)$ is the sample mean of $\log(m_{t}(x))$ over $t$. The log transformation introduces some bias, so the LC method adjusts $\beta_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $\beta_t$ is then extrapolated using a random walk with drift model, which is equivalent to an ARIMA ($0,1,0$) model with a drift parameter.

@@

By conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}$ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of $y_{n+h}(x)$ are given by
\begin{equation}
  \widehat{y}_{n+h\mid n}(x)=\widehat{s}_{n+h\mid n}(x) = \text{E}[\widehat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \widehat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\widehat{\beta}_{n+h\mid n,j},
\end{equation}
where $\widehat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because $\left\{\beta_{t,j}\right\}$ is time-varying, it controls the dynamics of the process. Each component series $\left\{\beta_{t,j}\right\}$ is modelled with an ARIMA$(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are determined by minimizing the Akaike Information Criterion [@forecastpackage], and the parameters are estimated using maximum likelihood estimation.

@@

When we apply the @HU07 model to data on males and females, we need to constrain the resulting forecasts so they do not diverge. In the product-ratio method [@HBY13], we apply the HU model to the geometric mean of male and female log mortality data, and the ratio of male and female log mortality data, rather than to each series independently. That is, we define
\begin{align}
  p_t(x)&=\sqrt{s_{t}^{\text{M}}(x)s_{t}^{\text{F}}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t}^{\text{M}}(x)}{s_{t}^{\text{F}}(x)}}. \label{eq:coherent_2}
\end{align}
Then, the HU method is applied to these two new variables instead of male and female series:
\begin{align*}
  \log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  \log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with an ARIMA model while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA model (i.e., the order of differencing is constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecasts of male and female do not diverge. Then, forecasts of sex-specific data are obtained using
\begin{align*}
  m_{n+h\mid n}^{\text{M}}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n}^{\text{F}}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for forecasting net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration, and a sum and difference is used rather than a product and ratio.

@@

Following the general procedure proposed in @HB08, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future.

We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration (whereas @HB08 used the HU method for all five components).

@@

It should be noted that there are three types of randomness in the simulation: randomness from the second equation in the Hyndman-Ullah specification (i.e., $e_t(x)$ in \eqref{eq:HU_1}), randomness from the ARIMA model fitted to principal component scores, and randomness from the Poisson distribution for births and deaths or the bootstrap for migrants.

@@

Historical central death rates and start-year and mid-year populations of residents in Australia by sex and age in single years for age group 0--99 and 100+ have been sourced from the @HMD for 1921--2018. The start-year population for 2019 is also used to estimate net migration. Data for age-specific annual fertility rates by single years of age for 15--49 over the period 1950--2018 was obtained from the @ABSBirths. Even though annual mortality rates are available from 1921 to 2018, only data collected after 1950 is used for two reasons. First, the period between 1921--1950 has different mortality patterns from later years because of wars and epidemics [@HB08]. Since the model for this study is valid under the assumption that extrapolative methods can be used, it seemed best to delete the less relevant data, which might influence estimations of parameters if we include them. Second, fertility rates are only available from 1950, so the mortality rates were also truncated to the same period.

@@

We use the fitted model of fertility rates as an example to interpret the results of fitted models. It should be noted that models fitted to product and ratio series are less able to be interpreted in an economic sense. Figure \@ref(fig:componentsfertility) shows the estimated mean age pattern $\widehat{\mu}(x)$, the first two basis functions, $\widehat{\phi}_1(x)$ and $\widehat{\phi}_2(x)$, and the first two principal component scores, $\widehat{\beta}_{t,1}$ and $\widehat{\beta}_{t,2}$) for fertility rates. The first term ($\widehat{\phi}_1(x) \widehat{\beta}_{t,1}$) accounts for the most variation in fertility. The first principal component scores $\widehat{\beta}_{t,1}$, shown in black for 1950--2018, display a sharp increase between 1950--1960, followed by a rapid decline in fertility rates from 1961. This increase of fertility is linked to the Post-World War II baby boom in Australia, resulting from the recovery from war and the Great Depression. The dramatic decrease from 1960 was due to the availability of contraception pills in Australia. A noticeable further reduction of fertility rates in the late 1960s may be due to a change in the law allowing abortion in some states in Australia.

@@

The first basis function $\widehat{\phi}_1(x)$ measures the different effects of the time-trend across ages. It indicates the general pattern of fertility has little effect on females of age 15, most likely because such pregnancies are almost always accidental and usually prior to the mother taking contraception. The two "bumps" in the first basic function suggests that the historical trend has a distinct impact on females aged around their 20s and 40s. The second basis function largely concerns mothers aged over 30. The forecasts of future coefficients are shown as the blue lines in the lower panels with their 80% pointwise prediction intervals shown as blue shaded regions. These forecasts are obtained using automatically selected ARIMA models. In practice, we use six basis functions, but the four not shown have a much smaller effect on the resulting fertility forecasts and are much harder to interpret.

@@

Figures \@ref(fig:totalpopulationbyyears) and \@ref(fig:totalpopulationin2034) show the forecast results of population. Figure \@ref(fig:totalpopulationbyyears) shows the forecast of the total population by sex over the prediction horizon, while Figure \@ref(fig:totalpopulationin2028) shows the forecast of population age structure in a specific year, 2038. Figure \@ref(fig:totalpopulationbyyears) indicates that both the female and male population are expected to increase to around 27 million by 2068. Since the fertility rate is expected to remain low, the possible reason for the significant increase in population size is net migration. We can also observe that the further the projected year is, the bigger the forecast variation of a population. This increase in variation shows that we have less relevant information for further apart prediction horizons.

```{r simulation, dependson='models'}
############### SIMULATION STEP #################
# This takes a while!
set.seed(2020 - 05 - 02)
nsim <- 20 # Change to 2000 or higher to get better estimates
aus.sim <- pop.sim(mort = mortf, fert = fertf, mig = migf, firstyearpop = aus.pop, N = nsim)

# Combine simulation results into one tibble
dimnames(aus.sim[["male"]]) <- dimnames(aus.sim[["female"]]) <- list(
  Age = dimnames(aus.sim[[1]])[[1]],
  Year = dimnames(aus.sim[[1]])[[2]],
  Rep = dimnames(aus.sim[[1]])[[3]]
)
aussim_df <- bind_rows(
  as.tbl_cube(aus.sim[["male"]] / 1e3, met_name = "Population") %>%
    as_tibble() %>% mutate(Sex = "Male"),
  as.tbl_cube(aus.sim[["female"]] / 1e3, met_name = "Population") %>%
    as_tibble() %>% mutate(Sex = "Female")
)
# rm(aus.sim)
# Combine simulations to create means and quantiles of future years
future_pop <- aussim_df %>%
  group_by(Sex, Age, Year) %>%
  summarise(
    Mean = mean(Population),
    Lower = quantile(Population, prob = 0.1),
    Upper = quantile(Population, prob = 0.9)
  ) %>%
  ungroup()
# Historical data as a tibble
history_pop <- bind_rows(
  as_tibble(aus.pop$pop$male / 1000) %>%
    mutate(
      Age = aus.pop$age,
      Sex = "Male"
    ) %>%
    pivot_longer(c(-Age, -Sex), names_to = "Year", values_to = "Population"),
  as_tibble(aus.pop$pop$female / 1000) %>%
    mutate(
      Age = aus.pop$age,
      Sex = "Female"
    ) %>%
    pivot_longer(c(-Age, -Sex), names_to = "Year", values_to = "Population"),
) %>%
  mutate(Year = as.numeric(Year))
```

```{r totalpopulationbyyears, fig.cap="Fifty-year forecasts of the total population for each sex, along with 80% prediction intervals. The actual population from 1950 to 2019 is shown using circles.", fig.height=4, dependson='simulationresults'}
future_tot_pop <- aussim_df %>%
  group_by(Rep, Sex, Year) %>%
  summarise(Population = sum(Population) / 1e3) %>%
  group_by(Sex, Year) %>%
  summarise(
    Mean = mean(Population),
    Median = median(Population),
    Lo95 = quantile(Population, prob = 0.025),
    Up95 = quantile(Population, prob = 0.975),
    Lo80 = quantile(Population, prob = 0.1),
    Up80 = quantile(Population, prob = 0.9)
  ) %>%
  ungroup()
history_pop %>%
  group_by(Sex, Year) %>%
  summarise(Population = sum(Population) / 1e3) %>%
  ggplot(aes(x = Year, y = Population, group = Sex)) +
  geom_line(size = 1) +
  ylab("Population (millions)") +
  ggtitle("Total population") +
  facet_grid(. ~ Sex) +
  geom_ribbon(aes(y = Mean, ymin = Lo80, ymax = Up80), fill = "#7d7def", data = future_tot_pop) +
  geom_line(data = future_tot_pop, col = "#0000cc", mapping = aes(y = Mean), size = 1)
```

```{r totalpopulationin2034, fig.cap="Forecast population pyramid for 2038, along with 80% pointwise prediction intervals. The actual pyramid for 2019 is shown using dashed lines.", warnings=FALSE, dependson='simulationresults'}
pyramid <- function(history, future) {
  data <- bind_rows(history, future)
  p_male <- data %>%
    filter(Sex == "Male") %>%
    ggplot(aes(x = Age, y = Population, group = Year)) +
    labs(y = "Population ('000)") +
    geom_ribbon(aes(x = Age, ymin = Lower, ymax = Upper, y = Mean), fill = "#7d7def") +
    geom_line(linetype = 2) +
    geom_line(aes(y = Mean), col = "#0000cc") +
    coord_flip() +
    scale_y_reverse(expand = c(0.012, 0.012)) +
    scale_x_continuous(expand = c(0.012, 0.012)) +
    facet_grid(. ~ Sex)
  p_female <- data %>%
    filter(Sex == "Female") %>%
    ggplot(aes(x = Age, y = Population, group = Year)) +
    labs(y = "Population ('000)") +
    geom_ribbon(aes(x = Age, ymin = Lower, ymax = Upper, y = Mean), fill = "#7d7def") +
    geom_line(linetype = 2) +
    geom_line(aes(y = Mean), col = "#0000cc") +
    coord_flip() +
    scale_y_continuous(expand = c(0.012, 0.012)) +
    theme(
      axis.title.y.left = element_blank(),
      axis.text.y.left = element_blank(),
      axis.ticks.y.left = element_blank()
    ) +
    scale_x_continuous(position = "top", sec.axis = sec_axis(~.), expand = c(0.012, 0.012)) +
    facet_grid(. ~ Sex)

@@

Figure \@ref(fig:totalpopulationin2034) shows the mean and the 80% pointwise prediction interval of simulated population paths for the year 2038, along with the 2019 base population. It can be observed that the overall population size is expected to increase substantially. The uncertainty of the infant population is much higher than other ages, which reflects the higher forecast variation of fertility rates than mortality rates. The uncertainty of the young age population is mainly due to migration [@HB08].

```{r OADRat65, fig.cap="Fifty-year forecasts of the OADR with the assumption of pension age equal to 65 for both sexes, along with 80% pointwise prediction interval.", dependson='simulationresults'}
OADR_history <- oadr(history_pop, filter(pension_age, Policy == "Current"))
OADR.65 <- oadr_sim(aussim_df, filter(pension_age, Policy == "Current")) %>%
  mutate(Policy = "Pension age 65")
OADR.65 %>%
  ggplot(aes(x = Year, y = OADR)) +
  geom_ribbon(aes(ymin = Lo, ymax = Hi), fill = "#7d7def") +
  geom_line(col = "#0000aa", size = .75) +
  geom_line(col = "black", size = .75, data = OADR_history) +
  xlab("year") +
  ylab("ratio") +
  ggtitle("OADR forecasts with pension age=65")
```

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the OADR associated with a pension age equal to 65 (green line) and equal to the pension age scheme with approved change (orange line) respectively, along with the desired OADR level in purple. The actual OADRs for 1921--2018 are shown in black.", dependson='OADRat65'}
df <- bind_rows(
  OADR.65,
  oadr_sim(aussim_df, filter(pension_age, Policy == "Approved")) %>%
    mutate(Policy = "Current policy"),
  oadr_sim(aussim_df, filter(pension_age, Policy == "Proposed")) %>%
    mutate(Policy = "Proposed policy")
) %>%
  mutate(Policy = factor(Policy,
    levels = c("Pension age 65", "Current policy", "Proposed policy")
  ))
df %>%
  ggplot(aes(x = Year, y = OADR)) +
  geom_hline(yintercept = 0.21, col = "gray") +
  geom_ribbon(aes(ymin = Lo, ymax = Hi, fill = Policy, group = Policy), alpha = 0.5) +
  geom_line(aes(col = Policy, group = Policy), size = .75) +
  geom_line(col = "black", size = .75, data = OADR_history) +
  xlab("year") +
  ylab("ratio") +
  ggtitle("OADR forecasts with pension age=65") +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2")
```

The analysis was performed by first generating an OADR associated with a fixed pension age of 65 and an OADR associated with pension age scheme with the approved change, gradually increasing the pension age from 65 to 67 by mid-2023. Even though the government might further increase the pension age in future, we kept the pension age at 67 from 2024 to see what would happen if no further action was taken. Figure \@ref(fig:OADRat65) shows the mean forecast and pointwise prediction interval of the OADR with a pension age of 65, along with the historical OADR for 1921--2018. A strong positive trend was observed, which indicates an increasing burden on the next working-age generation if the pension age is fixed at 65. This is consistent with the prediction by inter-generation reports in 2002 [@IGR02] and 2007 [@IGR07]. According to former Australian Treasurer @Costello04, this could be a result of better health care and low fertility rates. These two reasons could be regarded as permanent change and therefore a persistent trend. In addition, there is a big jump from actual data to predicted data. This jump is possibly a result of the Baby Boomers’ retirement. Those born in 1946, the start of the baby boom, would begin retiring in 2011 when the pension age was 65.

Figure \@ref(fig:OADRcurrent) presents a comparison of the mean forecast of OADR with the pension age equal to 65 and a pension age equal to the pension age scheme with the approved change, along with a comparison of both forecasts with the desired level of OADR, which has been set at 21%. It should be noted that OADR corresponding to pension age scheme with approved change is approaching the desired level, which suggests that the approved change does achieve the aim of providing an affordable financial burden, at least for a moment. However, once the pension age reaches 67, the OADR of the pension scheme with approved adjustment grows at the nearly same rate as the OADR of a pension age equal of 65. This is consistent with the idea that there is a persistent positive trend in OADR associated with a constant pension age, due to low fertility rates and better health care. Therefore, Figure \@ref(fig:OADRcurrent) demonstrates that the pension age scheme requires constant review to ensure a stable OADR over the long term. However, the questions of when the action needs to be taken and how much of an adjustment needs to be made remain to be answered. This answer will allow us to respond to the recent proposal of further increasing the pension age to 70 by 2035.

n our sample, $n=2018-1950+1=69$ and we set $k=25$ and $H=25$ to perform the test. The mean absolute error of $h$-step-ahead forecast of OADR for $h=1,2,\dots,25$ is shown in Figure \@ref(fig:rollingtest). From this Figure, we observe that the mean absolute error is increasing as $h$ increases, which is reasonable as we have less relevant information to make a forecast for further apart OADR. Overall, the mean absolute error is small. For example, the mean absolute error for 25-step-ahead forecast is 1.6%, which is small if compared to the desired level of OADR, which is at 21%.

>>>>

```{r changespension, echo=FALSE, fig.cap="Australian pension age policies.", fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Policies obtained from https://www.dss.gov.au/seniors/benefits-payments/age-pension
# Law: https://guides.dss.gov.au/guide-social-security-law/3/4/1/10
# History: https://nationalseniors.com.au/news/campaign/a-short-history-of-the-age-pension-section
pension_age <- tibble(
  Year = 1920:2068,
  Current = 65
) %>%
  mutate(
    Proposed = pmin(70, Current + 0.5 * pmax(0, trunc((Year - 2015) / 2))),
    Approved = pmin(Proposed, 67),
  ) %>%
  pivot_longer(-Year, values_to = "Age", names_to = "Policy")

pension_age %>%
  filter(Year >= 2008, Year <= 2050) %>%
  ggplot(aes(x = Year, y = Age, group = Policy, col = Policy)) +
  geom_line() +
  ylab("Minimum pension age") +
  scale_color_brewer(
    palette = "Dark2",
    breaks = c("Proposed", "Approved", "Current"),
    labels = c("Proposed further change", "Approved change", "Pension age at 65")
  )
```

@@

Variants of the Lee-Carter method [@LC92] are frequently used for demographic forecasting [see, e.g., @SBH11]. The most recent forecast of the population in Australia by the @PC13 was based on basic Lee-Carter methods plus some judgement-based adaptations to address the characteristics of demographic trends that the Lee-Carter method failed to capture. The Hyndman-Ullah method (2007) is a generalization of the Lee-Carter method and is used as the basis for forecasting demographic components in this study. In addition, the product-ratio method [@HBY13] is used to eliminate the potential long-term divergence between female and male mortality rates and net-migration rates.

@@

The model is over-parameterised. Therefore, the constraints
\begin{equation}
  \sum_{t=1}^{n}\beta_t=0, \qquad\text{and}\qquad \int_x \phi(x)=1
\end{equation}
are imposed, and a singular value decomposition is used to find the least squares solution of \eqref{eq:lc} [@Good1969]. The estimates of $\phi(x)$ and $\beta_t$ are therefore the first principal component and scores of $\bm{X}$ with $(i,t)$th element $[\log(m_{t}(x))-\widehat\mu(x)]$, where $\widehat\mu(x)$ is the sample mean of $\log(m_{t}(x))$ over $t$. The log transformation introduces some bias, so the LC method adjusts $\beta_t$ such that for each year, the implied number of deaths will be equal to the actual number of deaths. The adjusted $\beta_t$ is then extrapolated using a random walk with drift model.

@@

By conditioning on the observed data $\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}$ and the set of functional principal components $\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}$, the $h$-step-ahead forecast of $y_{n+h}(x)$ are given by
\begin{equation}
  \widehat{y}_{n+h\mid n}(x)=\widehat{s}_{n+h\mid n}(x) = \text{E}[\widehat{s}_{n+h\mid n}(x)\mid \bm{I, B}] = \widehat{\mu}(x)+\sum_{j=1}^{J}\phi_j(x)\widehat{\beta}_{n+h\mid n,j},
\end{equation}
where $\widehat{\beta}_{n+h\mid n,j}$ denotes the $h$-step-ahead forecast of $\beta_{n+h\mid n,j}$. Because $\left\{\beta_{t,j}\right\}$ is time-varying, it controls the dynamics of the process. Each component series $\left\{\beta_{t,j}\right\}$ is modelled with an ARIMA$(p_j,d_j,q_j)$ process, where $p_j,d_j,q_j$ are selected using the automated algorithm of @forecastpackage, and the parameters are estimated using maximum likelihood estimation.

@@

When we apply the @HU07 model to data on males and females, we need to constrain the resulting forecasts so they do not diverge. In the product-ratio method [@HBY13], the HU model is applied to the geometric mean of male and female log mortality data, and the ratio of male and female log mortality data, rather than to each series independently. That is, define
\begin{align}
  p_t(x)&=\sqrt{s_{t}^{\text{M}}(x)s_{t}^{\text{F}}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t}^{\text{M}}(x)}{s_{t}^{\text{F}}(x)}}. \label{eq:coherent_2}
\end{align}
Then, the HU method is applied to these two new variables:
\begin{align*}
  \log[p_t(x)]=\mu_p(x)+\sum_{j=1}^{J}\beta_{t,j}\phi_j(x)+e_t(x)\\
  \log[r_t(x)]=\mu_r(x)+\sum_{l=1}^{L}\gamma_{t,l}\psi_l(x)+z_t(x)
\end{align*}
where time series $\left\{\beta_{t,j}\right\}$ are fitted with ARIMA models while $\left\{\gamma_{t,l}\right\}$ are fitted by stationary ARFIMA models (i.e., the order of differencing is constrained to $0<d<\frac{1}{2}$). It is the stationary constraint on $\gamma_{t,l}$ that ensures the forecasts of male and female do not diverge. Then, forecasts of sex-specific data are obtained using
\begin{align*}
  m_{n+h\mid n}^{\text{M}}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n}^{\text{F}}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}
The method for forecasting net migration by sex is similar to the above method used on mortality rates except no log scale is applied to net migration, and a sum and difference is used rather than a product and ratio.

@@

Following the general procedure proposed in @HB08, we model each of the five components of population change: female mortality rates, male mortality rates, female net migration, male net migration and fertility rates, and simulate these five components into the future. We use the HU method applied to log-fertility, and the product-ratio method applied to log-mortality and net-migration.

@@

There are three types of randomness in the simulation: (1) randomness from the second equation in the Hyndman-Ullah specification (i.e., $e_t(x)$ in \eqref{eq:HU_1}); (2) randomness from the ARIMA model fitted to principal component scores; and (3) randomness from the Poisson distribution for births and deaths, or the bootstrap for migrants.

@@

Historical central death rates and start-year and mid-year populations of residents in Australia by sex and age in single years for age group 0--99 and 100+ have been sourced from the @HMD for 1921--2018. The start-year population for 2019 is also used to estimate net migration. Data for age-specific annual fertility rates by single years of age for 15--49 over the period 1950--2018 was obtained from the @ABSBirths. Even though annual mortality rates are available from 1921 to 2018, only data collected after 1950 is used for two reasons. First, the period between 1921--1950 has different mortality patterns from later years because of wars [@HB08]. Since the model for this study is valid under the assumption that extrapolative methods can be used, it seemed best to delete the less relevant data, which might influence estimations of parameters if we include them. Second, fertility rates are only available from 1950, so the mortality rates were also truncated to the same period.

@@

We illustrate the component model forecasts using fertility rates. Figure \@ref(fig:componentsfertility) shows the estimated mean age pattern $\widehat{\mu}(x)$, the first two basis functions, $\widehat{\phi}_1(x)$ and $\widehat{\phi}_2(x)$, and the first two principal component scores, $\widehat{\beta}_{t,1}$ and $\widehat{\beta}_{t,2}$) for log fertility rates. The first term, $\widehat{\phi}_1(x) \widehat{\beta}_{t,1}$, accounts for `r round(fert.fit$varprop[1]*100,1)`% of the variation in the log fertility rates. The first principal component scores $\widehat{\beta}_{t,1}$, shown in black for 1950--2018, display a sharp increase between 1950--1960, followed by a rapid decline in fertility rates from 1961. This increase of fertility is linked to the post-World War II baby boom, while the dramatic decrease from 1960 was due to the availability of contraception pills. A noticeable further reduction of fertility rates in the late 1960s may be due to a change in the abortion laws in some states.

@@

The first basis function $\widehat{\phi}_1(x)$ measures the different effects of the time-trend across ages. It indicates the general pattern of fertility has little effect on females of age 15, most likely because such pregnancies are almost always accidental and usually prior to the mother taking contraception. The two "bumps" in the first basic function suggests that the historical trend has a distinct impact on females aged around their 20s and 40s. The second basis function, accounting for `r round(fert.fit$varprop[2]*100,1)`% of the variation, largely concerns mothers aged over 30. The forecasts of future coefficients are shown as the blue lines in the lower panels with their 80% pointwise prediction intervals shown as blue shaded regions. These forecasts are obtained using automatically selected ARIMA models. In practice, we use six basis functions, but the four not shown have a much smaller effect on the resulting fertility forecasts, only account for `r round(sum(fert.fit$varprop[3:6])*100,1)`% of the variation, and are much harder to interpret.

@@

Figure \@ref(fig:totalpopulationbyyears) shows the forecast of the total population by sex over the prediction horizon, while Figure \@ref(fig:totalpopulationin2038) shows the 20-year forecast of population age structure in 2038. Figure \@ref(fig:totalpopulationbyyears) indicates that both the female and male population are expected to increase to around 27 million by 2068.

```{r simulation, dependson='models'}
############### SIMULATION STEP #################
if(file.exists("aus_sim.rds")) {
  aus.sim <- readRDS("aus_sim.rds")
} else {
  # This takes a while!
  set.seed(2020 - 05 - 02)
  nsim <- 2000 # Change to 2000 or higher to get better estimates
  aus.sim <- pop.sim(mort = mortf, fert = fertf, mig = migf, firstyearpop = aus.pop, N = nsim)
  saveRDS(aus.sim, "aus_sim.rds")
}
```

```{r sim_tibbles, dependson='simulation'}
# Combine simulation results into one tibble
dimnames(aus.sim[["male"]]) <- dimnames(aus.sim[["female"]]) <- list(
  Age = dimnames(aus.sim[[1]])[[1]],
  Year = as.numeric(dimnames(aus.sim[[1]])[[2]]) + 1,
  Rep = dimnames(aus.sim[[1]])[[3]]
)
aussim_df <- bind_rows(
    as.tbl_cube(aus.sim[["male"]] / 1e3, met_name = "Population") %>%
      as_tibble() %>% mutate(Sex = "Male"),
    as.tbl_cube(aus.sim[["female"]] / 1e3, met_name = "Population") %>%
      as_tibble() %>% mutate(Sex = "Female")
  )

# Combine simulations to create means and quantiles of future years
future_pop <- aussim_df %>%
  group_by(Sex, Age, Year) %>%
  summarise(
    Mean = mean(Population),
    Lower = quantile(Population, prob = 0.1),
    Upper = quantile(Population, prob = 0.9)
  ) %>%
  ungroup()
# Historical data as a tibble
history_pop <- bind_rows(
  as_tibble(aus.pop$pop$male / 1000) %>%
    mutate(
      Age = aus.pop$age,
      Sex = "Male"
    ) %>%
    pivot_longer(c(-Age, -Sex), names_to = "Year", values_to = "Population"),
  as_tibble(aus.pop$pop$female / 1000) %>%
    mutate(
      Age = aus.pop$age,
      Sex = "Female"
    ) %>%
    pivot_longer(c(-Age, -Sex), names_to = "Year", values_to = "Population"),
) %>%
  mutate(Year = as.numeric(Year))
```

```{r totalpopulationbyyears, fig.cap="Fifty-year forecasts of the total population for each sex, along with 80% prediction intervals.", fig.height=4, dependson='simulationresults'}
future_tot_pop <- aussim_df %>%
  group_by(Rep, Sex, Year) %>%
  summarise(Population = sum(Population) / 1e3) %>%
  group_by(Sex, Year) %>%
  summarise(
    Mean = mean(Population),
    Median = median(Population),
    Lo95 = quantile(Population, prob = 0.025),
    Up95 = quantile(Population, prob = 0.975),
    Lo80 = quantile(Population, prob = 0.1),
    Up80 = quantile(Population, prob = 0.9)
  ) %>%
  ungroup()
history_pop %>%
  group_by(Sex, Year) %>%
  summarise(Population = sum(Population) / 1e3) %>%
  ggplot(aes(x = Year, y = Population, group = Sex)) +
  geom_line(size = 1) +
  ylab("Population (millions)") +
  ggtitle("Total population") +
  facet_grid(. ~ Sex) +
  geom_ribbon(aes(y = Mean, ymin = Lo80, ymax = Up80), fill = "#7d7def", data = future_tot_pop) +
  geom_line(data = future_tot_pop, col = "#0000cc", mapping = aes(y = Mean), size = 1)
```

```{r totalpopulationin2038, fig.cap="Forecast population pyramid for 2038, along with 80% pointwise prediction intervals. The actual pyramid for 2019 is shown using dashed lines.", warnings=FALSE, dependson='simulationresults'}
pyramid <- function(history, future) {
  data <- bind_rows(history, future)
  p_male <- data %>%
    filter(Sex == "Male") %>%
    ggplot(aes(x = Age, y = Population, group = Year)) +
    labs(y = "Population ('000)") +
    geom_ribbon(aes(x = Age, ymin = Lower, ymax = Upper, y = Mean), fill = "#7d7def") +
    geom_line(linetype = 2) +
    geom_line(aes(y = Mean), col = "#0000cc") +
    coord_flip() +
    scale_y_reverse(expand = c(0.012, 0.012)) +
    scale_x_continuous(expand = c(0.012, 0.012)) +
    facet_grid(. ~ Sex)
  p_female <- data %>%
    filter(Sex == "Female") %>%
    ggplot(aes(x = Age, y = Population, group = Year)) +
    labs(y = "Population ('000)") +
    geom_ribbon(aes(x = Age, ymin = Lower, ymax = Upper, y = Mean), fill = "#7d7def") +
    geom_line(linetype = 2) +
    geom_line(aes(y = Mean), col = "#0000cc") +
    coord_flip() +
    scale_y_continuous(expand = c(0.012, 0.012)) +
    theme(
      axis.title.y.left = element_blank(),
      axis.text.y.left = element_blank(),
      axis.ticks.y.left = element_blank()
    ) +
    scale_x_continuous(position = "top", sec.axis = sec_axis(~.), expand = c(0.012, 0.012)) +
    facet_grid(. ~ Sex)

@@

Figure \@ref(fig:totalpopulationin2038) shows the mean and the 80% pointwise prediction interval of simulated population paths for the year 2038 (a 20-year forecast), along with the 2019 base population. The uncertainty of the population below aged 20 is largely due to higher forecast variation of fertility rates compared to other components.

Historical OADR values are shown in Figure \@ref(fig:OADRcurrent) for 1921--2019. The strong positive trend indicates an increasing tax burden on workers, consistent with the prediction of the intergenerational reports in 2002 [@IGR02] and 2007 [@IGR07]. The increased trend from 2010 is due to the retirements of "Baby Boomers", born after WW2. We also show forecasts for three pension schemes: (1) a fixed pension age of 65; (2) the current scheme with the pension age rising to age 67 in 2023; and (3) the proposed pension policy under which the pension ages continues to rise 6 months every two years until it reaches 70 in 2035.

```{r OADRcurrent, fig.cap="Fifty-year forecasts of the OADR associated with a pension age equal to 65 (green), the current approved pension age scheme (orange), and the proposed scheme with the pension age rising to 70 in 2035 (purple). The actual OADRs for 1921--2019 are shown in black.", dependson='OADRcurrent'}
OADR_history <- oadr(history_pop, filter(pension_age, Policy == "Current"))
OADR.65 <- oadr_sim(aussim_df, filter(pension_age, Policy == "Current")) %>%
  mutate(Policy = "Pension age 65")
df <- bind_rows(
  OADR.65,
  oadr_sim(aussim_df, filter(pension_age, Policy == "Approved")) %>%
    mutate(Policy = "Current policy"),
  oadr_sim(aussim_df, filter(pension_age, Policy == "Proposed")) %>%
    mutate(Policy = "Proposed policy")
) %>%
  mutate(Policy = factor(Policy,
    levels = c("Pension age 65", "Current policy", "Proposed policy")
  ))
df %>%
  ggplot(aes(x = Year, y = OADR)) +
  geom_hline(yintercept = 0.25, col = "gray") +
  geom_ribbon(aes(ymin = Lo, ymax = Hi, fill = Policy, group = Policy), alpha = 0.5) +
  geom_line(aes(col = Policy, group = Policy), size = .75) +
  geom_line(col = "black", size = .75, data = OADR_history) +
  xlab("year") +
  ylab("ratio") +
  ggtitle("OADR forecasts with pension age=65") +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2")
```

Once the pension age reaches 67, the OADR under the current policy grows at nearly the same rate as the OADR of a pension age equal of 65. This is consistent with the idea that there is a persistent positive trend in OADR associated with a constant pension age, due to low fertility rates and better health care. Even the proposal to raise the pension age incrementally to 70 by 2035 leads to a similar increasing trend beyond 2035. Figure \@ref(fig:OADRcurrent) demonstrates that the pension age scheme requires constant review to ensure a stable OADR over the long term.
